---
title: Exploratory analysis for life expectancy and other elements of the life table,
  Canada, all provinces except Prince Edward Island dataset
author: "Eva Y"
date: '2018-10-12'
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

knitr::opts_knit$set(root.dir = "../")
```

```{r load dependency}
library(data.table)
library(ggplot2)
```

```{r load data}
life_data <- fread("data/data_hackseq/13100114.csv", sep = ",")
metadata <- fread("data/data_hackseq/13100114_MetaData.csv", sep = ",")
```

## Smell Test Data

### What is the data that we have?

```{r Smell test, echo=FALSE}
names(life_data)
dim(life_data)
```

There are 17 columns and here's what I think they are. 

- REF_DATE: Between different years
- GEO: Geographic location
- DGUID: ?
- Age group: Remove 'years' and replace the x in Element with years in age group. 
- Sex: Simplify to F, M, or Both
- Element: Remove spaces in between
- UOM: ?
- UOM_ID: ?
- SCALAR_FACTOR: ?
- SCALAR_ID: ?
- VECTOR: ?
- COORDINATE: ?
- VALUE: Most important
- STATUS: ?
- SYMBOL: ?
- TERMINATED: ?
- DECIMALS: ?

```{r Clean data}
# change column names
setnames(life_data, old = "Age group", new = "Age_group")

# remove unimportant columns
life_data[, c("DGUID", "UOM", "UOM_ID", "SCALAR_FACTOR", "SCALAR_ID", "VECTOR", "COORDINATE", "STATUS", "SYMBOL", "TERMINATED", "DECIMALS") := NULL]

# clean values in each column
life_data[, Age_group := gsub("[[:space:]].*", "", Age_group)]
life_data[, Sex := gsub()]
```

```{r}
life_data[GEO == "Canada", unique(Element)]
```

