---
title: Exploratory analysis for life expectancy and other elements of the life table,
  Canada, all provinces except Prince Edward Island dataset
author: "Eva Y"
date: '2018-10-12'
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

knitr::opts_knit$set(root.dir = "../")
```

```{r load dependency}
library(data.table)
library(ggplot2)
```

```{r load data}
life_data <- fread("data/data_hackseq/13100114.csv", sep = ",")
metadata <- fread("data/data_hackseq/13100114_MetaData.csv", sep = ",")
```

## Smell Test Data

### What is the data that we have?

```{r Smell test, echo=FALSE}
names(life_data)
dim(life_data)
```

There are 17 columns and here's what I think they are. 

- REF_DATE: Between different years
- GEO: Geographic location
- DGUID: ?
- Age group: Remove 'years' and replace the x in Element with years in age group. 
- Sex: Simplify to F, M, or Both
- Element: Remove spaces in between
- UOM: ?
- UOM_ID: ?
- SCALAR_FACTOR: ?
- SCALAR_ID: ?
- VECTOR: ?
- COORDINATE: ?
- VALUE: Most important
- STATUS: ?
- SYMBOL: ?
- TERMINATED: ?
- DECIMALS: ?

```{r Clean data}
# change column names
setnames(life_data, old = "Age group", new = "Age_group")

# remove unimportant columns
life_data[, c("DGUID", "UOM", "UOM_ID", "SCALAR_FACTOR", "SCALAR_ID", "VECTOR", "COORDINATE", "STATUS", "SYMBOL", "TERMINATED", "DECIMALS") := NULL]

# clean values in each column
life_data[, Age_group := gsub("[[:space:]].*", "", Age_group)]
life_data[, Sex := gsub("Males", "M", Sex)]
life_data[, Sex := gsub("Females", "F", Sex)]
life_data[, Sex := gsub("Both sexes", "Both", Sex)]

# change Value to numeric
life_data[, VALUE := as.numeric(VALUE)]

# split REF_DATE column
life_data[, c("START_YR", "END_YR") := tstrsplit(REF_DATE, "/", fixed = T)]
```

### How does clean data looks like? 

```{r View clean data}
head(life_data)
```

## Create dataset with separate years

```{r Create dataset with separate years, echo=FALSE}
# for each REF_DATE, I would take the start year and end year and create a column of in between years, then merge. This will give me a list of tables, which I will merge at the end.

create_sep_year <- function(ref_date, dt){

  # create small table with the years
  dt <- dt[REF_DATE == ref_date]
  interval <- seq(dt[, unique(START_YR)], dt[, unique(END_YR)])
  # dummy_dt <- data.table(YEAR = rep(interval, times = nrow(dt)))
  
  # create table with each row repeated for the number of years in one interval
  dt <- dt[rep(1:.N, each = length(interval))]
  big_dt <- cbind(interval, dt)
}

ref_date <- life_data[, unique(REF_DATE)]

# testing code
# test <- tail(life_data)
# ref_date <- test[, unique(REF_DATE)]
# test2 <- lapply(ref_date, create_sep_year, dt = test)
# test2 <- rbindlist(test2)

# run function
life_data_sep <- lapply(ref_date, create_sep_year, dt = life_data)
life_data_sep <- rbindlist(life_data_sep)

# change column name
setnames(life_data_sep, old = "interval", new = "YEAR")
```

```{r Check out data}
head(life_data_sep)
```

> How many rows and columns are there? 

```{r Dimensions for clean data}
dim(life_data_sep)
```

Number of rows is `r nrow(life_data_sep)` and number of columns is `r ncol(life_data_sep)`. This seems correct since the original data has `r nrow(life_data)` and each interval covers three years. 

## Average duplicated years

To separate the data based on individual years, we created duplicates for years that are within the overlapping year intervals. Now, the task is to get an average for these duplicates. 

```{r average the years based on year}
life_data_sep_avg <- life_data_sep[, .(AVG_VALUE = mean(VALUE)), by = c("YEAR", "GEO", "Age_group", "Sex", "Element")][order(YEAR)]
head(life_data_sep_avg)
```

```{r Save csv, eval=FALSE}
write.table(life_data_sep_avg, "results/life_data_sep_avg.csv", sep = ",")
```

